Even-point Multi-loop Unitarity and its Applications: Exponentiation, Anomalies and Evanescence

### 2023-07-07 eventPointTwoLoop.tex
corrected 4.87 state sum
fixed labels and state sum in 4.86 and 87

Luckily, the state-sewing for maximal supersymmetry is dramatically more simple than less than maximal supersymmetry \cite{BDKUniarityReview,BRY}. This is most easily seen by considering the covariant operators we defined for one-loop DBIVA amplitudes. When one applies the conditions for maximal supersymmetry, stated below,
\begin{align}
D=10 \qquad N_\lambda &=1 \qquad N_X=0
\\
D=6 \qquad N_\lambda &=2 \qquad N_X=1
\\
D=4 \qquad N_\lambda &=4 \qquad N_X=3
\\
D=3 \qquad N_\lambda &=8 \qquad N_X=8
\end{align}
then the state sum is {completely independent of loop momenta} and precisely reproduces the supersymmetric operator $s_{12}^2 (t_8F^4)$ when cut along the $s_{12}$-channel.  This statement holds covariantly in general dimension. Concretely, we find that for theories with maximal supersymmetry:
\begin{equation}
\sum_{\text{states}}(t_8F^4)^{(\text{max})}_{(12l_1l_2)}(t_8F^4)^{(\text{max})}_{(\bar{l}_1\bar{l}_234)} = s_{12}^2 (t_8F^4)^{(\text{max})}_{(1234)}
\end{equation}
This can similarly be used in the integrand construction $\mathcal{N}=4$ super-Yang-Mills, which has been computed to six-loop order \cite{Carrasco:2021otn}. In the case of maximal supersymmetry in $D=4$, we can write the  $(t_8F^4)^{(\text{max})}_{(1234)}$ operator as a supersymmetric delta function with the all-plus permutation invariant introduced previously:
\begin{equation}
\mathcal{M}^{\mathcal{N}=4\text{DBIVA}}_{(1234)} = \delta^{(8)}{(Q)} \frac{[12][34]}{\langle 12\rangle \langle 34\rangle} \equiv (t_8F^4)^{(\text{max})}_{(1234)}\Big|_{D=4}
\end{equation}
where the delta function is a Grassmann valued polynomial of spinor-helicity variables:
\begin{equation}
\delta^{(8)}{(Q)} = \prod_{a=1}^4\sum_{i\neq j} \langle ij \rangle \eta_i^a\eta_j^a
\end{equation}
By applying this supersymmetric state sum, we find that the integrands for two-loop $\mathcal{N}=4$ DBIVA is trivially easy to construct - even simpler than NLSM integrands. Below we represent an internal on-shell superfield with a green line, and obtain the following integrand for the double-bubble:
\begin{equation}\label{eq:N42bub}
\text{Cut}\left(\scaleIntCtune{}{hgreen}{}{hgreen}\right)=\sum_{\text{states}}  (t_8F^4)^{(\text{max})}_{(1234)} (t_8F^4)^{(\text{max})}_{(1234)} (t_8F^4)^{(\text{max})}_{(1234)}= s_{12}^4 (t_8 F^4)
\end{equation}
and similarly so for the ostrich-diagram integral contribution:
\begin{equation}\label{eq:N4ostrich}
\text{Cut}\left(\scaleIntBtune{}{}{}{}{hgreen}{hgreen}{hgreen}{hgreen}\,\,\right)=\sum_{\text{states}}  (t_8F^4)^{(\text{max})}_{(1234)} (t_8F^4)^{(\text{max})}_{(1234)} (t_8F^4)^{(\text{max})}_{(1234)}=s_{12}^2 \tau^{(2)}_2\tau^{(1)}_1(t_8 F^4)\,.
\end{equation}
\dj{Relabel actual cuts and lets check the RHS here of the ostrich.  I get: $s_{12}^2 \left(\tau^{(1)}_{\bar{2}} \equiv (k_1-q_2)^2 \right)^2 (t_8 F^4)$ both via KLT on the cuts and via your sewing rule.}

We should point out an additional verification that can occur for any SUSY amount of DBIVA.  We can recycle multiloop SYM integrands to verify our cuts.  In general this is because double-copy works at tree-level, and we can always write the Jacobi-satisfying Yang-Mills cut numerators in terms of color-ordered sYM amplitudes, $A^{\text{sYM}}$, offering us a KLT between cut-integrands.  When we cut to four-point amplitudes, as we do for the cuts above, this is particularly simple as only one order needs contribute for each for-point tree, yielding:
\begin{align}
\text{Cut}({\text{DBIVA}})&= \sum_\text{states} \prod_i \mathcal{A}^{\text{DBIVA}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \sum_\text{states} \prod_i  \left( s^{(i)}_{12} s^{(i)}_{23}  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\right)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right)  \right) \sum_\text{states}   \prod_i  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right) \right) \times \text{Cut}({\text{sYM}})\,.
\end{align}
Here  the products run over the number of trees contributing to the cut,  $s^{(i)}_{mn} = ( p^{(i)}_m + p^{(i)}_n )^2$, and $p^{(i)}_m$ is the $m$-th label of the $i$-th tree contributing to the cut. The second line follows from the double-copy structure of four-point amplitudes in DBIVA, where:
\begin{align}
\mathcal{A}_4^{\text{DBIVA}} = s_{12} \, s_{23} \, A^{\text{sYM}}(1,2,3,4)\,.
\end{align}

>>>>

Luckily, the state-sewing for maximal supersymmetry is dramatically more simple than less than maximal supersymmetry \cite{BDKUniarityReview,BRY}. This is most easily seen by considering the covariant operators we defined for one-loop DBIVA amplitudes. When one applies the conditions for maximal supersymmetry, stated below,
D=10 \qquad N_\lambda &=1 \qquad N_X=0
D=6 \qquad N_\lambda &=2 \qquad N_X=1
D=4 \qquad N_\lambda &=4 \qquad N_X=3
D=3 \qquad N_\lambda &=8 \qquad N_X=8
then the state sum is {completely independent of loop momenta} and precisely reproduces the supersymmetric operator $s_{12}^2 (t_8F^4)$ when cut along the $s_{12}$-channel.  This statement holds covariantly in general dimension. Concretely, we find that for theories with maximal supersymmetry:
\sum_{\text{states}}(t_8F^4)^{(\text{max})}_{(12l_1l_2)}(t_8F^4)^{(\text{max})}_{(\bar{l}_1\bar{l}_234)} = s_{12}^2 (t_8F^4)^{(\text{max})}_{(1234)}
This can similarly be used in the integrand construction $\mathcal{N}=4$ super-Yang-Mills, which has been computed to six-loop order \cite{Carrasco:2021otn}. In the case of maximal supersymmetry in $D=4$, we can write the  $(t_8F^4)^{(\text{max})}_{(1234)}$ operator as a supersymmetric delta function with the all-plus permutation invariant introduced previously:
\mathcal{M}^{\mathcal{N}=4\text{DBIVA}}_{(1234)} = \delta^{(8)}{(Q)} \frac{[12][34]}{\langle 12\rangle \langle 34\rangle} \equiv (t_8F^4)^{(\text{max})}_{(1234)}\Big|_{D=4}
where the delta function is a Grassmann valued polynomial of spinor-helicity variables:
\delta^{(8)}{(Q)} = \prod_{a=1}^4\sum_{i\neq j} \langle ij \rangle \eta_i^a\eta_j^a
By applying this supersymmetric state sum, we find that the integrands for two-loop $\mathcal{N}=4$ DBIVA is trivially easy to construct - even simpler than NLSM integrands. Below we represent an internal on-shell superfield with a green line, and obtain the following integrand for the double-bubble:
\begin{equation}\label{eq:N42bub}
\text{Cut}\left(\scaleIntCtune{}{hgreen}{}{hgreen}\right)=\sum_{\text{states}}  (t_8F^4)^{(\text{max})}_{(p_4 12p_3)} (t_8F^4)^{(\text{max})}_{(\bar{p}_1\bar{p}_4\bar{p}_3\bar{p}_2)} (t_8F^4)^{(\text{max})}_{(p_234p_1)}= s_{12}^4 (t_8 F^4)
and similarly so for the ostrich-diagram integral contribution:
\begin{equation}\label{eq:N4ostrich}
\text{Cut}\left(\scaleIntBtune{}{}{}{}{hgreen}{hgreen}{hgreen}{hgreen}\,\,\right)=\sum_{\text{states}}  (t_8F^4)^{(\text{max})}_{(2q_4\bar{q}_3\bar{q}_1)} (t_8F^4)^{(\text{max})}_{(1\bar{q}_4q_3\bar{q}_2)} (t_8F^4)^{(\text{max})}_{(q_234q_1)}=s_{12}^2 \tau^{(1)}_{\bar{2}}\tau^{(2)}_{\bar{1}}(t_8 F^4)\,.
We should point out an additional verification that can occur for any SUSY amount of DBIVA.  We can recycle multiloop SYM integrands to verify our cuts.  In general this is because double-copy works at tree-level, and we can always write the Jacobi-satisfying Yang-Mills cut numerators in terms of color-ordered sYM amplitudes, $A^{\text{sYM}}$, offering us a KLT between cut-integrands.  When we cut to four-point amplitudes, as we do for the cuts above, this is particularly simple as only one order needs contribute for each for-point tree, yielding:
\text{Cut}({\text{DBIVA}})&= \sum_\text{states} \prod_i \mathcal{A}^{\text{DBIVA}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \sum_\text{states} \prod_i  \left( s^{(i)}_{12} s^{(i)}_{23}  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\right)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right)  \right) \sum_\text{states}   \prod_i  A^{\text{sYM}}(p^{(i)}_1,p^{(i)}_2,p^{(i)}_3,p^{(i)}_4)\,\\
&= \left( \prod_j  \left(s^{(j)}_{12} s^{(j)}_{23}\right) \right) \times \text{Cut}({\text{sYM}})\,.
Here  the products run over the number of trees contributing to the cut,  $s^{(i)}_{mn} = ( p^{(i)}_m + p^{(i)}_n )^2$, and $p^{(i)}_m$ is the $m$-th label of the $i$-th tree contributing to the cut. The second line follows from the double-copy structure of four-point amplitudes in DBIVA, where:
\mathcal{A}_4^{\text{DBIVA}} = s_{12} \, s_{23} \, A^{\text{sYM}}(1,2,3,4)\,.
